---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aquaman

<!-- badges: start -->
[![R-CMD-check](https://github.com/aquaMetrics/aquaman/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/aquaMetrics/aquaman/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/aquaMetrics/aquaman/branch/main/graph/badge.svg)](https://app.codecov.io/gh/aquaMetrics/aquaman?branch=main)
<!-- badges: end -->

### WORK IN PROGRESS - DO NOT USE IN PRODUCTION

The goal of `aquaman` R package is to calculate ecology metrics and assess the mixing zone
area, fulfilling these specific steps:

 - Assign bacteria taxa from DNA reads using `dada2` package.
 - Model IQI metric from assigned bacteria families.
 - Assess the mixing zone size from IQI sampling results.

## Installation

You can install the development version of aquaman like so:

```r
install.packages("devtools")
devtools::install_github("aquaMetrics/aquaman")
```

## Area Assessment

This basic example shows you how to calculate mixing zone from IQI
input data:

```{r example}
library(aquaman)
```

```r
## Run the demo IQI data and return the area of the mixing zone
area <- assess(demo_iqi)
area$fifthPercentileArea
#>      5% 
#> 96914.92 
```

## Model IQI

Using assigned family-level bacteria taxa as predictors, calculate benthic
invert IQI as an outcome.

```{r}
# Run the IQI model based on demo taxanomic data
iqi_scores <- iqi(demo_taxa)
head(iqi_scores, 5)
```

IQI prediction model created by Tom Wilding (SAMS) based on training data from
SEPA And MOWI.

## Assign Taxa

Assign bacteria families based on S-16 DNA reads. This in time may provide the
input to `iqi()` function. Currently, a Qiime 2 command line script achieves
this part of the process.

*Note*, you will be prompted to download a reference taxonomic file on first
use. This will be stored locally.

```r
# Provide a path to the demo data within the package:
taxa <- assign_taxa(demo_path())
# ...this could take some time...
head(taxa, 5)
#>        sample_id                 Family reads
#> 1   MHS-ARD-0-E2           Mitochondria     8
#> 2   MHS-ARD-0-E2           Mitochondria    47
#> 3   MHS-ARD-0-E2           Mitochondria   384
#> 4   MHS-ARD-0-E2        Anaerolineaceae     7
#> 5   MHS-ARD-0-E2                   <NA>    10
```

