---
title: "qiime2 pipeline"
output: html_document
vignette: >
  %\VignetteIndexEntry{qiime2}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
library(aquaman)
library(kraken)
```

## Assign Taxa

Run the qiime2 bash file on MacOS or Linux. Recommend Windows Subsystem for Linux (WSL) if on Windows machine. This script requires the installation of qiime2 and conda libraries.

Additionally, need to update script to provide a path to directory containing .fastq files from Illumina instrument.

```{r run_qiime2}

path = "C:\\Users\\Tim.Foster\\OneDrive - Scottish Environment Protection Agency\\Reports\\Fish Farm\\operating_iqi_surveys\\ScreeningTool_SequenceData\\Site12R2AA_fastq"

path <- demo_path()

script_path <- system.file("extdat",
    "qiime2-bash-script", package = "aquaman")
# Only on Linux, Mac or WSL
system(paste0("bash ", script_path, "/qiime2-taxa-assign.txt ", path))



```

## Import taxa

```{r import}

taxa <- import_qiime2(path = "")

```

## Calculate IQI

```{r}

iqi_score <- iqi(taxa)

```

## Import Survey

```{r}

survey <- import_survey(path = "")


```

## Prepare

Merge taxa, survey, iqi scores.

```{r}
data <- inner_join(survey, iqi_score, taxa)
```

## Calculate Area

```{r}

area <- assess(data)

```
